---
title: "Machine Learning Course Project"
author: "Zelepukin Dmitriy"
date: "12 august 2017"
output:
  html_document: default
  pdf_document: default
---
The goal of this project is to predict the manner in which they did the exercise. This is the "classe" variable in the training set. 

#Load libraries and data sets
```{r setup, include=FALSE}
library(rpart)
library(rattle)
library(caret)
library(randomForest) # random forest
train <- read.csv('pml-training.csv', stringsAsFactors=F, na.strings = c("NA", "", "#DIV/0!"))
test <- read.csv('pml-testing.csv', stringsAsFactors=F, na.strings = c("NA", "", "#DIV/0!"))
```

# There is need to prepare data for a random forest method
```{r precleaning}
train <- train[,-c(1:7)]
table(train$classe)

new.names <- c("classe")

for(i in names(train)[1:152]){
    
    #--- step 1
    share.na <- round(sum(is.na(train[,i]))*100/19622,2)
    print(paste0(i, " has NAs: ", share.na, "%"))
    
    #--- step 2
    if(share.na==0){
        new.names<-c(new.names,i)
    }
}

new.train <- train[,new.names]
new.train$classe <- as.factor(new.train$classe)
table(new.train$classe)
```

Random Forest method prediction - training and cross-validation checking: 
---------------
```{r randforest}
set.seed(1234)
cv3 = trainControl(method="cv",number=3,allowParallel=TRUE,verboseIter=TRUE)
modrf = train(classe~., data=new.train, method="rf",trControl=cv3)
```

And add a decision tree model:
```{r dec_tree}
modtree = train(classe~.,data=new.train,method="rpart",trControl=cv3)
prf = predict(modrf,new.train)
ptree = predict(modtree,new.train)
```

Let`s check models:
```{r check}
table(prf,new.train$classe)
table(ptree,new.train$classe)
prf=predict(modrf,test)
ptree=predict(modtree,test)
table(prf,ptree)
```

Finally, the random forest model runs with a better accuracy.